---
title: "Statistical Inference Assignment Week 3- Part 2"
author: "Nidhi Shrivastava"
date: "Friday, March 20, 2015"
output: word_document
---
- Load the ToothGrowth data and perform some basic exploratory data analyses 
- Provide a basic summary of the data.
- Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class,                   even if there's other approaches worth considering)
- State your conclusions and the assumptions needed for your conclusions. 

**Q1 . Load the ToothGrowth data and perform some basic exploratory data analyses.**


```{r, echo=FALSE}
library(ggplot2)
library(datasets)
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
plot <- ggplot(ToothGrowth, 
               aes(x=factor(dose),y=len,fill=factor(dose)))
plot + geom_boxplot(notch=F) + facet_grid(.~supp) +
        scale_x_discrete("Dosage (Milligram)") +   
        scale_y_continuous("Length of Teeth") +  
        ggtitle("Basic Exploratory Data Analyses")

plot <- ggplot(ToothGrowth, 
               aes(x=factor(supp),y=len,fill=factor(supp)))
plot + geom_boxplot(notch=F) + facet_grid(.~dose) +
        scale_x_discrete("Supplement") +   
        scale_y_continuous("Length of Teeth") +  
        ggtitle("Data Analyses- Tooth length with different supplement for each dosage")
```

**Q2 . Provide a basic summary of the data.**
```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
summary(ToothGrowth)
table(ToothGrowth$supp, ToothGrowth$dose)
```

**Q3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering).**


```{r, echo=FALSE}
supp.t1 <- t.test(len~supp, paired=F, var.equal=T, data=ToothGrowth)
supp.t2 <- t.test(len~supp, paired=F, var.equal=F, data=ToothGrowth)
supp.result <- data.frame("p-value"=c(supp.t1$p.value, supp.t2$p.value),
                          "Conf-Low"=c(supp.t1$conf[1],supp.t2$conf[1]),
                          "Conf-High"=c(supp.t1$conf[2],supp.t2$conf[2]),
                          row.names=c("Equal Var","Unequal Var"))
supp.result
```

**Q4. State your conclusions and the assumptions needed for your conclusions.**

Based on the analysis above, we can conclude that

 - The 2mg dose has larger impact on tooth growth than 1mg and 0.5mg, while 1mg dose has more impact than 0.5mg dose. 
 So there is a different in the growth of the tooth while the doses are larger.

 - There is no doubt that orange juice and vitamin C have obvious different impact on tooth growth.
