---
title: "Statistical Inference Course Project -part 1"
author: "Nidhi Shrivastava"
date: "Wednesday, March 18, 2015"
output: word_document
---


- Synopsis
- Task
- Question 1
       - Answer 1
- Question 2
       - Answer 2
- Question 3
       - Answer 3



**Synopsis**

```
This is the project for Statstical Inference. We have to create the simulation exercises for exponential distribution to theoretical distribution.

The project consists of two parts:
- A simulation Exercise
- Basic Inferential data analysis
```
**Task**

```
The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. 

In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. Note that you will need to do a thousand or so simulated averages of 40 exponentials.

```
```
## set seed for reporducibility
set.seed(3)
```
```

## set lambda to 0.2

lambda<-0.2
```
```

## 1000 simulations

nsims<-1000
```
```

## 40 exponential samples
n<-40
```
```

## see the number of simulation in a matrix form

sim<-matrix(rexp(nsims*n, rate=lambda), nsims, n)

```
```

## calculate means

row_means<-rowMeans(sim)

```
```
Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s. You should:
```
**Question 1**

```
1. Show the sample mean and compare it to the theoretical mean of the distribution.

```
```

#distribution mean

theoretical_mean<-mean(row_means)
theoretical_mean
## [1] 4.987876
```
```
#sample mean

 sample_mean<-1/lambda
 sample_mean
## [1] 5
```
```
#visualization - draw a histogram

hist(row_means, breaks=50, prob=TRUE, main="Distribution of averages of samples, drawn from exponential distribution with lambda=0.2", xlab="")

```
```
## density of the averages sample

lines(density(row_means))
```
```
##theoretical center of distribution

abline(v=1/lambda, col='red')
```
```
##theoretical density of averages of samples
xfit<-seq(min(row_means), max(row_means), length=100)
```
```
yfit<-dnorm(xfit, mean=1/lambda, sd=((1/lambda)/sqrt(n)))
```
```
lines(xfit, yfit, pch=22, col="red", lty=2)
```
```
##add legend
legend('topright', legend=c("simulation","theoretical"), lty=1:2, col("black", "red"))




```

```{r, echo=FALSE}
set.seed(3)
lambda<-0.2
nsims<-1000
n<-40
## see the number of simulation in a matrix form

sim<-matrix(rexp(nsims*n, rate=lambda), nsims, n)

row_means<-rowMeans(sim)

hist(row_means, breaks=50, prob=TRUE, main="Distribution of averages of samples", xlab="")

## density of the averages sample

lines(density(row_means))
##theoretical center of distribution

abline(v=1/lambda, col='red')

##theoretical density of averages of samples

xfit<-seq(min(row_means), max(row_means), length=100)

yfit<-dnorm(xfit, mean=1/lambda, sd=((1/lambda)/sqrt(n)))

lines(xfit, yfit, pch=22, col="red", lty=2)


##add legend
legend('topright', legend=c("simulation","theoretical"), lty=1:2, col("black", "red"))

```

**Answer 1**

The analytics mean is 4.993867 the theoretical mean 5. The center of distribution of averages of 40 exponentials is very close to the theoretical center of the distribution.


**Question 2**

Show how variable it is and compare it to theoretical variance of the distribution.

```
##standard deviation of distribution
s<-sd(row_means)
s
## [1] 0.7834318

```

```
##standard deviation from the sample 

s1<- (1/lambda)/sqrt(n)
 s1
##[1] 0.7905694

```
```
##variance of distribution

var_dist<-s^2
 var_dist
## [1] 0.6137653
```
```
##variance of theoretical mean
var_theory<-((1/lambda)*(1/sqrt(n)))^2
var_theory
## [1] 0.625
```



**Answer 2**
Standard Deviation of the distribution is 0.7834318 with the theoretical SD calculated as 0.7905694. The Theoretical variance is calculated as ($(\frac{1}{lambda} * \frac{1}{sqrt(n)})^2$) = 0.625. The actual variance of the distribution is 0.6137653.



**Question 3**
Show that the distribution is approximately normal.


```
xfit<-seq(min(row_means), max(row_means), length=100)

yfit<-dnorm(xfit, mean=1/lambda, sd=((1/lambda)/sqrt(n)))

hist(row_means, breaks=n, prob=TRUE, col="orange", xlab="Means", main="Density of Means", ylab="Density")

lines(xfit, yfit, pch=22, col="red", lty=2)

## add legend
legend('topright', legend=c("simulation","theoretical"), lty=1:2, col("black", "red"))

```


```{r echo=FALSE}


##theoretical density of averages of samples

xfit<-seq(min(row_means), max(row_means), length=100)

yfit<-dnorm(xfit, mean=1/lambda, sd=((1/lambda)/sqrt(n)))

hist(row_means, breaks=n, prob=TRUE, col="orange", xlab="Means", main="Density of Means", ylab="Density")

lines(xfit, yfit, pch=22, col="red", lty=2)


##add legend
legend('topright', legend=c("simulation","theoretical"), lty=1:2, col("black", "red"))

```


```
##compare the distribution of averages of 40 exponentials to a normal distribution

qqnorm(row_means)

qqline(row_means, col = 2)


```


```{r, echo=FALSE}


qqnorm(row_means)

qqline(row_means, col = 2)

```


**Answer 3**

Due to the central limit theorem (CLT), the distribution of averages of 40 exponentials is very close to a normal distribution.
